Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-12-16 PAPER/Paola_2/ms_ver3.tex
Corrected fig9 and fig12

\subsection{SiO abundance}
\label{ssec:sioabun}
Molecular abundance is one of the important ingredients that is
required by the radiative transfer code described above. 
Typically, extremely low abundance of SiO is
found (n(SiO)/n(H2) $< 3\times10^{-12}$) in dark, dense clouds such
as TMC1 \citep{Ziurys:1989p14699,MartinPintado:1992p14309}. Whereas, in
outflows like L1448, SiO abundance can increase up to 10$^{-6}$ especially in
molecular bullets moving with a projected velocity of 60
km\,s$^{-1}$ \citep{MartinPintado:1992p14309, Dutrey:1997p11185}. Thus, there is
an increase of 6 orders of magnitude from quiescent clouds
to outflows. Production of
gaseous SiO due to slow C-type shocks has been suggested to occur via release of
silicon from grain cores and from grain mantles. Various stationary
shock models indicate a sudden abundance increase in SiO 
near a shock speed of 20-30\,km\, s$^{-1}$. However, several young
outflows have velocities of the order of 100 km\,s$^{-1}$ as in the present
case. Shocks due to such outflows will dissociate H$_{2}$ and will
become J ({\em{jump}}) type shocks. Thus, molecules observed in such
energetic outflows must have been reformed in the flow as suggested by
detailed models of J-shocks by \cite{Neufeld:1989p14322}. SiO formation in J shocks
has also been modelled recently and has thought to be the source of
the SiO line emission observed in molecular outflows and jets
\citep{Guillet:2009p11229}.
%

In-spite of all numerical models relating to the study of enhancement
of SiO in shocks, very little is known about the dependence of SiO
abundance on shock speeds. Considering the complex grain
chemistry that is involved in order to estimate the functional
dependence of SiO abundance on shock speeds, we prescribe these
profiles based on limited empirical evidences. The most simple among
them is the top hat profile in which the SiO abundance is a low value of
10$^{-12}$ below 20 km s$^{-1}$ and above 100 km s$^{-1}$ and a
maximum value of 10$^{-6}$ between these two velocities.
Shock velocities $>$ 100 km s$^{-1}$ can lead to the 
ionisation of hydrogen and thereby effectively reduce the number
of collision centers for SiO resulting in a weaker emission. 
However, we also consider a step function dependence of SiO abundance
on shock velocity to incorporate cases where we allow for reformation
of atomic and molecular species.
In order to get rid of a discontinuous change of abundance, we also
prescribe a gaussian such that the peak SiO abundance of 10$^{-6}$ is at
60\,km\,s$^{-1}$, while the value of n(SiO)/n(H$_{2}$) falls below
10$^{-9}$ at 20 and 100 km s$^{-1}$. Such a huge variation in SiO
abundance may indeed be the case in regions where SiO is seen
in widespread \citep[e.g.,][also see section~\ref{sec:discussion}]
{Lefloch:1998p13983}. Also along with a Gaussian profile, we consider
the case where we extent the abundance of 10$^{-6}$ beyond 60 km\,s$^{-1}$. 
These functional dependences of SiO
abundance on shock velocity are shown in figure~\ref{abun}.  
In addition to this, if the temperature at the point is greater than
92,000\,K (the temperature of the Si-O bond disassociation energy) 
the abundance is reduced to 0. 
%

 \begin{figure}
  \includegraphics[width=1.\columnwidth]{\figpath/pfig3.pdf}%{\figpath/diffabun_profnew.png}
  \caption{The different fractional SiO abundances as a function of
    jet velocity, v$_{\rm jet}$.}
  \label{abun}
 \end{figure}

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the present study, we will consider two
cases. Firstly, the most simple one in which the jet flow velocity is
same as the shock velocity. Considering that there is a density
contrast in the simulation setup, this assumption will
overestimate the shock velocity. We also
consider a case in which the shock speed of a dense plug accumulated
between the jet and medium is estimated by balancing the respective ram
pressures \citep{Masson:1993p9661},
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm jet}}{1 + \eta^{-0.5}} = v_{\rm jet}\delta
\end{equation}
where v$_{\rm s}$ is the velocity of dense plug and v$_{\rm jet}$ is the
jet velocity. A comparative view of using such velocity corrections and
functional forms of SiO abundances are discussed in
Sec~\ref{ssec:sioabunem}.

% The fractional abundance is given by the equation:

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig12_temp.pdf}%almamaps_new.pdf}%{\figpath/ALMAmaps3.png}
 \caption{{\bf Left:} The integrated intensity map of SiO(2-1), (5-4)
   and (8-7). The emission map shows the J=5-4 line intensity (in units
   of Jy$\,$kms$^{-1}$/beam), the blue contours show the J=8-7 line
   intensity and the green contours show the J=2-1 line intensity. The
   jet is inclined at an angle of 60$^{\circ}${\bf Right:} The PV
   diagram taken along the axis of the jet for the 5-4 line (in units
   of Jy/beam), showing the higher J transitions highlighting the
   knots of the jet and broad emission at the bow shock. 
   In both panels the ticks on the color bar represent the different contour levels.}
\label{fig:almafig}
\end{figure*}

>>>>

\subsection{SiO abundance}
\label{ssec:sioabun}
Molecular abundance is one of the important ingredients that is
required by the radiative transfer code described above. 
Typically, extremely low abundance of SiO is
found (n(SiO)/n(H2) $< 3\times10^{-12}$) in dark, dense clouds such
as TMC1 \citep{Ziurys:1989p14699,MartinPintado:1992p14309}. Whereas, in
outflows like L1448, SiO abundance can increase up to 10$^{-6}$ especially in
molecular bullets moving with a projected velocity of 60
km\,s$^{-1}$ \citep{MartinPintado:1992p14309, Dutrey:1997p11185}. Thus, there is
an increase of 6 orders of magnitude from quiescent clouds
to outflows. Production of
gaseous SiO due to slow C-type shocks has been suggested to occur via release of
silicon from grain cores and from grain mantles. Various stationary
shock models indicate a sudden abundance increase in SiO 
near a shock speed of 20-30\,km\, s$^{-1}$. However, several young
outflows have velocities of the order of 100 km\,s$^{-1}$ as in the present
case. In case of L1448, the typical velocities of working surfaces (internal knots) 
lie within a range of 80-150 km\,s$^{-1}$ and the SiO abundance 
lie around 10$^{-7}$ to 10$^{-6}$ for such high velocity bullets \citep{Dutrey:1997p11185}. 
Shocks due to such outflows could dissociate H$_{2}$ and molecules observed in such
energetic outflows could have been reformed in the flow as suggested by
detailed models of J-shocks by \cite{Neufeld:1989p14322}. SiO formation in J shocks
has also been modelled recently and has thought to be the source of
the SiO line emission observed in molecular outflows and jets
\citep{Guillet:2009p11229}.
%

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the bow shock, one can estimate the shock velocity 
by equating the pressures across the shock surface, \citep[e.g.,][]{Masson:1992p14101}
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm jet}}{1 + \eta^{-0.5}}
\end{equation}
where v$_{\rm s}$ is the velocity of the bow shock 
and v$_{\rm jet}$ is the jet velocity. 
Considering the dependence of ambient density on the vertical distance, 
the density contrast at the bow shock increases to $\eta \sim 100$. 
Thus, for such large $\eta$ values, one can very well 
approximate the bow shock velocity to be the same as the jet flow velocity.  
In case of internal knots, the above equation for the velocity of the working surface
changes to eq.~\ref{eq:velws} \citep{Raga:1992p16392}. 
As described in the appendix ~\ref{sec:worksurface}, the velocity of the internal 
knots effectively reduces to an average of upstream and downstream 
velocities across the knot and can as well be approximated 
to be same as flow velocity. However, the shock velocity 
in the frame of reference of the knot can be considerably different than the flow velocity 
and is expressed as a difference between the upstream and downstream velocities.  
In the present models, these internal shock velocities are $\leq$ 20 km\,s$^{-1}$. 
Steady state calculations of shocked grain 
micro-physics described above, takes into account 
these internal shock velocities to estimate 
SiO abundances. While, observations measure bulk velocities of the 
flow and relate molecular abundances to the same. 
%

In-spite of all numerical models relating to the study of enhancement
of SiO in shocks, very little is known about the dependence of SiO
abundance on shock speeds. Considering the complex grain
chemistry that is involved in order to estimate the functional
dependence of SiO abundance on shock speeds, we prescribe these
profiles based on limited empirical evidences 
and approximate the variation of SiO abundances on jet flow velocities.   
The most simple and extreme choice among
them is the step profile in which the SiO abundance is a low value of
10$^{-12}$ below 20 km\,s$^{-1}$ and a maximum value of 10$^{-6}$ beyond. 
Such a choice effectively reduces to assuming a constant value of high abundance for the 
flow velocities seen in our model. In order to compare such a choice, 
we also apply a top-hat profile where above 100 km\,s$^{-1}$, abundance value 
is set again to a low value of 10$^{-12}$.
In order to get rid of an arbitrary discontinuous change of abundance, we also
prescribe a Gaussian such that the peak SiO abundance of 10$^{-6}$ is at
60\,km\,s$^{-1}$, while the value of n(SiO)/n(H$_{2}$) falls below
10$^{-9}$ at 20 and 100 km s$^{-1}$. Such a huge variation in SiO
abundance may indeed be the case in regions where SiO is seen
in widespread \citep[e.g.,][also see section~\ref{sec:discussion}]
{Lefloch:1998p13983}. Also along with a Gaussian profile, we consider
the case where we extent the abundance of 10$^{-6}$ beyond 60 km\,s$^{-1}$.   
In addition to this, if the temperature at the point is greater than
92,000\,K (the temperature of the Si-O bond disassociation energy) 
the abundance is reduced to 0. 
These functional dependence of SiO abundance on jet flow velocities are shown in figure~\ref{abun} 
and a comparative view is further discussed in Sec~\ref{ssec:sioabunem}.
%

 \begin{figure}
  \includegraphics[width=1.\columnwidth]{\figpath/pfig3.pdf}%{\figpath/diffabun_profnew.png}
  \caption{The different fractional SiO abundances as a function of
    jet velocity, v$_{\rm jet}$.}
  \label{abun}
 \end{figure}
% The fractional abundance is given by the equation:

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig12.pdf}%almamaps_new.pdf}%{\figpath/ALMAmaps3.png}
 \caption{{\bf Left:} The integrated intensity map of SiO(2-1), (5-4)
   and (8-7). The emission map shows the J=5-4 line intensity (in units
   of Jy$\,$kms$^{-1}$/beam), the blue contours show the J=8-7 line
   intensity and the green contours show the J=2-1 line intensity. The
   jet is inclined at an angle of 60$^{\circ}${\bf Right:} The PV
   diagram taken along the axis of the jet for the 5-4 line (in units
   of Jy/beam), showing the higher J transitions highlighting the
   knots of the jet and broad emission at the bow shock. 
   In both panels the ticks on the color bar represent the different contour levels.}
\label{fig:almafig}
\end{figure*}
